---
output: 
 md_document:
   variant: markdown_github
---
The purpose of this README is to guide the reader through the decisions I made when writing my final Financial Econometrics research assignment (Masters in Economics). I selected the following question:

* "Study bond yield spreads - you have 10yr and 2yr bond yields. Calculate the spread and see if you can find any interesting patterns comparing yield spreads through time, linking it to VIX / currency volatility - and perhaps looking at other macro factors too."

This README will guide the reader through any code, functions, and output that was constructed for this assignment. 

# Interpretation of the Question

```{r include=FALSE, warning=FALSE}

rm(list = ls())

gc()

###

library(tidyverse)
library(ggplot2)

```

# Required Packages:

* tidyverse
* ggplot2

# Loading and Editing Data sets

```{r}

  # 2 year Global Bonds.

Global_2yr <- readRDS("data/BondYields_2Y.rds") %>%
 
     mutate(Name = gsub("_2yr", "", Name)) %>% 
    
     filter(date <= "2022-03-31") # last available date in SA_2Yr

  # 10 year Global bonds.

Global_10yr <- readRDS("data/BondYields_10Y.rds") %>% 
    
    mutate(Name = gsub("_10Yr", "", Name)) %>% 
    
    filter(Name %in% Global_2yr$Name) %>% # Same bonds only
    
    filter(date <= "2022-03-31")

  # 2 year South African (SA) bonds. 

SA_2yr <- readRDS("data/BondYields_2Y_SA.rds") %>% 
    
    filter(date >= first(Global_10yr$date))

  # 10 Year SA bonds.

SA_10yr <- readRDS("data/BondYields_10Y_SA.rds") %>% 
    
    filter(date >= first(Global_10yr$date) & date <= "2022-03-31") 

```

The data sets all now start at the same date, and the global sets include the same bonds. At this point it is clear that we have COVID-19 Data. Perhaps we can do a sort of crisis bond yield analysis? 

```{r}

```

