---
output: 
 md_document:
   variant: markdown_github
---
The purpose of this README is to guide the reader through the decisions I made when writing my final Financial Econometrics research assignment (Masters in Economics). I selected the following question:

* "Study bond yield spreads - you have 10yr and 2yr bond yields. Calculate the spread and see if you can find any interesting patterns comparing yield spreads through time, linking it to VIX / currency volatility - and perhaps looking at other macro factors too."

This README will guide the reader through any code, functions, and output that was constructed for this assignment. 

# Interpretation of the Question

I have decided to focus on a comparison between South African bond yields and those of the rest of the world. I begin with descriptive data analysis, such as simply plotting the various yields - except I contextualize these yields by indicating high volatility times for the US and above average VIX periods. I also compare the distribution of yields for each country through the use of ridge lines. Finally, a visually intuitive tool - geospatial analysis - is used to convey they risk level of different countries. 

```{r include=FALSE, warning=FALSE}

rm(list = ls())

gc()

# General:

library(tidyverse)
library(ggplot2)
library(gridExtra)

# Ridge Plots:

library(ggridges)
library(viridis)
library(hrbrthemes)

# Geo-spatial data:

library(lwgeom)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

```

# Required Packages:

## General Packages

* tidyverse

## Plotting Packages

* ggplot2
* gridExtra

* ggridges 
* viridis 
* hrbrthemes 

* lwgeom 
* sf
* rnaturalearth 
* rnaturalearthdata

# Loading and Editing Data sets

## Importing Perscribed Data

The data sets all required similar wrangling, so I wrote a function to sort of get each data set into similar shape (Data_Creator_foo). The function essentially removes unnecessary tags from country names or bond tickers (so that I can more easily merge all sets later on), and converts to monthly yields. The reason for this is that I could only get monthly inflation from the World Bank and in case I want to do some yield adjustments later on, this change might be useful. Finally, the function filters only by dates post-2010, because the SA data sets were limited to this. 

```{r}


source("code/Data_Creator_foo.R")

# 2 year Global Bonds.

Global_2yr <- 

    Data_Creator_foo(readRDS("data/BondYields_2Y.rds"),
                 
                     "_2yr",
                 
                     "2Y")

  # 10 year Global bonds.

Global_10yr <- 
    
    Data_Creator_foo(readRDS("data/BondYields_10Y.rds"),
                     
                     "_10Yr",
                     
                     "10Y")  %>%
   
   filter(Name %in% Global_2yr$Name)



  # 2 year South African (SA) bonds. 

SA_2yr <- 
    
    Data_Creator_foo(readRDS("data/BondYields_2Y_SA.rds"),
                     
                     "_2Yr", 
                     
                     "")

  # 10 Year SA bonds.

SA_10yr <- 
    
    Data_Creator_foo(readRDS("data/BondYields_10Y_SA.rds"),
                     
                     "_10Yr",
                     
                     "") 

```

The code chunk below simply demonstrates my investigation into why the data sets might include the same names and bonds now, but do not have the same dimensions. This need not usually be a huge concern, because plotting will work regardless, but I just wanted to check whether there is cause for concern. In the end, it was fine and nothing much really needed changing. In order to check the similarity of different columns in each data set, I wrote a comparison function named Similarity_Checker_foo.R. 

```{r }

# Why are the data sets different lengths? 

source("code/Similarity_Checker_foo.R")

SimilarityChecker_foo(Global_10yr, Global_2yr, "date") # same dates 
SimilarityChecker_foo(Global_10yr, Global_2yr, "Name") # same countries 

SimilarityChecker_foo(
    
    Global_10yr, Global_2yr, "Ticker"
    
    ) # different bonds but same countries? 

###

Global_2yr <- 
    
    Global_2yr %>% 
    
    filter(
        
        Ticker %in% Global_10yr$Ticker) %>% 

# Removed some but still different number of observations. 
    
    #Duplicate dates? 

 distinct(Ticker, date, .keep_all = TRUE) # no duplicate dates 

# print(table(Global_10yr$Name))

# print(table(Global_2yr$Name))

# Different countries have different starting points in terms of dates, different numbers of observations each, between the two data sets. 

```

## Merging Perscribed Data 

Because the data came in pseudo-long format, I had to write a function to easily merge future data sets, as, from my point of view, merging these data sets required some substantial but repeatable coding, wrangling and thinking. So the function now does all of this for me - Data_Merger_foo.R. 

```{r }

source("code/Data_Merger_foo.R")

###

Bonds2yr <- 

Data_Merger_foo(Global_2yr, SA_2yr, BondYield_2, "BondYield_2")

###

Bonds10yr <- 

Data_Merger_foo(Global_10yr, SA_10yr, BondYield_10, "BondYield_10")

###

rm(list = c("Global_10yr", "Global_2yr", "SA_10yr", "SA_2yr"))

```

## Search Vectors 

Below, I simply create some search/character vectors that will easily allow me to filter through the data and only plot specific regions later on. This will help for the ridge line plots, and the bond yield plots - even the inflation plots to come. Region specific analysis makes more sense to me because one can otherwise get lost in a mess of lines or distributions that exhibit less than usual (visual) co-movement, and it can be difficult to see exactly what is going on. 

```{r }

# Isolating BRCIS countries: 

BRICS <- c(
    
    "Brazil", 
          
    "CHINA", 
          
    "India", 
           
    "Russia",
    
    "ZA"
    
    )

# Developed Countries: 

Dev_Count <- c(
    
    "Canada",
    
    "France",
    
    "Germany",
    
    "Japan",
    
    "Korea",
    
    "NZ",
    
    "Singapore",
    
    "UK",
    
    "US"
    
    )

# Developing Countries: 

Developing_vec <- 
    
    unique(
  
        Bonds2yr$Name[!grepl("EURO", Bonds2yr$Name) & 
                          
                        !grepl("Venezuela", Bonds2yr$Name) &
                          
                          !Bonds2yr$Name %in% Dev_Count]
        
)

# Venezuela destroyed the aesthetics of all my graphs because their yield, inflation and distribution are all ridiculous. It was visually easier to cope if they are not in the ridge or line plots. 

```

## Non-perscribed Data 

Below I download CPI and GDP Growth data from the World Bank. This is so that I am able to either use CPI, inflation, or GDP measures from a reputable source to contextualize bond yields. Inflation is one of the only pure economic risks that can face a bond - in normal times - and as such I think it is worth including in the analysis. I created a function to import the data from the World Bank as it is always in the same, messy and unclean format. So World_Bank_foo() gets it into the format I want. This was useful as I downloaded many data sets. 

I have also downloaded VIX and US-EU exchange rate data from FRED. With this VIX and US exchange rate data, I will not really be plotting them in a traditional way. They will be used to extract high volatility times so that I can add them to the line graphs as geom_rect() objects so as to be a contextual backdrop for the inflation and yield movements we observe. This will aid the discussion.

```{r }

# Search vector

All_Names <- unique(Bonds2yr$Name)

###

source("code/World_Bank_foo.R")

# World Bank growth rates: 

Growth_Rates <- 
    
    World_Bank_foo(read.csv("data/GDP Growth.csv"), # import data
                   
                   All_Names, # only these countries 
                   
                   Growth) # what to name gathered column


# world bank CPI measure, I have only downloaded since 2010. 

CPI <- 
    
  World_Bank_foo(read.csv("data/Inflation Data.csv"), All_Names, CPI) %>% 
    
    # I also create an inflation column: 
    
  arrange(Name, date) %>%
    
  group_by(Name) %>%
    
  mutate(Inflation = (CPI - lag(CPI)) / lag(CPI) * 100) %>% # inflation
    
  filter(date > "2009-01-01") %>% # first date is zero inflation 
    
  mutate(CPI = (CPI - 100)) %>% 
    
  ungroup()

###

```

```{r }

###

# VIX monthly data from FRED:

VIX_Data <- read.csv("data/VIX Data.csv") %>%
  
    mutate(date = as.Date(date, format = "%d/%m/%Y")) %>% 
    
        filter(date <= "2022-03-31")

###

US_Exchange <- read.csv("data/US to EU Exchange.csv") %>%
  
    mutate(date = as.Date(date, format = "%d/%m/%Y")) %>% 
    
        filter(date <= "2022-03-31")

###

source("code/High_Vol_foo.R")

VIX_Vol <- High_Vol_foo(VIX_Data, VIX, 0.8, type = "")

US_Vol <- High_Vol_foo(US_Exchange, Rate, 0.8, type = "sd")

# These high volatility periods are used to construct the geom_rects to indicate them on the ggplot graphs.

###

# I tried for hours to think of a functional way to create these rectangles. i.e. I thought maybe a for loop of some kind, i.e. for date in VIX_Vol etc... but I found it incredibly difficult to add whatever result I came up with neatly to a ggplot as an object. There probably is a very easy way to do it but it stumped me for quite some time. 

source("code/Rectangle_Creation_Script.R")

alpha = 0.15

```

# Bond Yields & Inflation 

Below, I plot the 10-year bond yields of three distinct regions. First, I plot the developed world, which exhibits considerable similarity in behavior. I also plot the developing world, and BRICS. Below each Bond yield plot, I also plot the inflation of each country. 

In these plots, light red zones are high VIX level time periods. The blue zones are high US exchange rate volatility time periods. If a zone is purple, it means these two zones overlap. In a zone such as this, one would expect considerably unique behavior - either from yields or from inflation.

The function created for this is named Yield_Plotter_foo. This is a poor name, in fact, because it can actually plot pretty much any variable that is in a tidy data set and requires a line graph to be drawn. It is also used to plot the inflation in each case, for example. 

## Developed Nations Bond Yields & Inflation 

```{r warning = FALSE, dpi=300, out.height= "100%", out.width= "100%"}

source("code/Yield_Plotter_foo.R")

###

grid.arrange(

Yield_Plotter(Data = Bonds10yr,
              Vec = Dev_Count, 
              y = BondYield_10,
              "Developed Nations",
              "Bond Yield (10-Year)",
              "none"),

Yield_Plotter(Data = CPI,
              Vec = Dev_Count, 
              y = Inflation,
              "",
              "Inflation",
              "bottom"),

ncol = 1,

heights = c(1.5,2.1))

```

## Developing Nations Bond Yields & Inflation 

```{r warning = FALSE, dpi=300, out.height= "100%", out.width= "100%"}

grid.arrange(

Yield_Plotter(Data = Bonds10yr,
              Vec = Developing_vec, 
              y = BondYield_10,
              "Developing Nations",
              "Bond Yield (10-Year)",
              "none"),

Yield_Plotter(Data = CPI,
              Vec = Developing_vec, 
              y = Inflation,
              "",
              "Inflation",
              "bottom"),

ncol = 1,

heights = c(1.5,2.1))


```

## BRICS Bond Yields & Inflation

```{r warning = FALSE, dpi=300, out.height= "100%", out.width= "100%"}

grid.arrange(

Yield_Plotter(Data = Bonds10yr,
              Vec = BRICS, 
              y = BondYield_10,
              "BRICS",
              "Bond Yield (10-Year)",
              "none"),

Yield_Plotter(Data = CPI,
              Vec = BRICS, 
              y = Inflation,
              "",
              "Inflation",
              "bottom"),

ncol = 1,

heights = c(2,2.5))

```

```{r}

# Housekeeping 

rm(
    
    list = c("HV1", "HV2", "HV3", "HV4", "HV5", "HV6", "HV7", "HV8")
    
    )

rm(
    
    list = c("VIX_Data", "US_Exchange", "US_Vol", "VIX_Vol")
    
    )

rm(list = "alpha")

```

## Ridge Plots 
  
Here, I simply plot the distribution of bond yields for all countries in the developed world, and BRICS - in a ridge line setup. The function created for this is Ridge_Plotter_foo. The dotted line represents, in the case of the developed world, the mean bond yield of the USA. In the case of the BRICS nations, it represents the mean bond yield of South Africa. 

```{r warning = FALSE, dpi=300, out.height= "100%", out.width= "100%"}

source("code/Ridge_Plotter_foo.R")

###

Ridge_Plotter(Bonds10yr, 
              "ZA", # South African mean bond yield dotted line
              BondYield_10,
              BRICS,
              "Distribution of BRICS Yields")

```

```{r warning = FALSE, dpi=300, out.height= "100%", out.width= "100%"}

Ridge_Plotter(Bonds10yr,
              "US", # American mean bond yield dotted line
              BondYield_10, 
              Dev_Count,
              "Distribution of Developed Countries Yields")

```

## Geospatial Plotting 

With geo-spatial plotting, the general issue that most would have is to get the data into the correct format. It must be an sf object, but you can download these from the geo-spatial packages in my package selection from earlier. Inspect the Map_Data_foo carefully, it simply takes the necessary steps in a repeatable process to merge the data sets in this case to the format required. 

```{r warning = FALSE, dpi=300, out.height= "100%", out.width= "100%"}

source("code/Map_Data_foo.R")

M1 <- Map_Data_foo(Bonds2yr, BondYield_2)
M2 <- Map_Data_foo(CPI, Inflation)
M3 <- Map_Data_foo(Growth_Rates, Growth)

```

Once the data is in the correct format, plotting simply requires ggplot. This function can do that, and does so repeatably. The big step is that it requires the geom_sf() notation in the plot, but other than that, it is plotted as usual and can be customized in identical ways. 

This geo-spatial plot simply demonstrates that being categorized by bond yield as high risk is not to be equated with being high risk due to inflation. There is definitely correlation, but the shade of countries is not identical in the two scenarios. 

```{r warning = FALSE, dpi=300, out.height= "100%", out.width= "100%"}

source("code/Create_Map_foo.R")

###

grid.arrange(

create_map(M1, "Average Bond Yield (2-Year)", "none"),

create_map(M2, "Average Inflation", "none"),

ncol = 1)

```

```{r warning = FALSE, dpi=300, out.height= "100%", out.width= "100%"}

create_map(M3, "Average GDP Growth", "none")

```

```{r}

rm(list = c("M1", "M2", "M3"))

```


# Yield Spreads

In order to plot the yield spreads (to answer the main question), I created Spread_Plotter(). It is a function that takes a data set, subtracts the US yield from the yield of all other countries on all dates, and then plots the resulting spread with a horizontal line at zero to indicate the US-zero level.

```{r warning = FALSE, dpi=300, out.height= "100%", out.width= "100%"}

source("code/Spread_Plotter.R")

###

grid.arrange(

   Spread_Plotter(Bonds2yr, 
                  BondYield_2, 
                  Dev_Count, 
                  "BondYield_2", 
                  "Developed Countries", 
                  "2-Year", 
                  "none"), 

   Spread_Plotter(Bonds10yr, 
                  BondYield_10, 
                  Dev_Count, 
                  "BondYield_10", 
                  "", 
                  "10-Year", 
                  "bottom"),

   ncol = 1,

   heights = c(1.5,2))

```

```{r warning = FALSE, dpi=300, out.height= "100%", out.width= "100%"}

grid.arrange(

   Spread_Plotter(Bonds2yr, 
                  BondYield_2, 
                  BRICS, 
                  "BondYield_2",
                  "BRICS", 
                  "2-Year", 
                  "none"), 

   Spread_Plotter(Bonds10yr, 
                  BondYield_10, 
                  BRICS, 
                  "BondYield_10", 
                  "", 
                  "10-Year", 
                  "bottom"),

   ncol = 1,

   heights = c(1.5,2))

```

